Bayesian model comparison for rare variant association studies of multiple phenotypes

### 2018-01-30 tex/manuscript.tex
Updates

\subsection*{Description of MRP}
In this section, we provide an overview of the MRP model comparison approach. Refer to \nameref{S1_Appendix} for detailed derivations. MRP models GWAS summary statistics as being distributed according to one of two models. The null model is that the regression effect sizes obtained across all studies for a group of variants and a group of phenotypes is zero. The alternative model is that summary statistics are distributed according to a multivariate normal distribution with mean zero and covariance matrix described below. MRP compares the evidence for the null and alternative model using a Bayes Factor (BF) that quantifies the amount of evidence for each model as the ratio of the marginal likelihoods of the observed data under two models.

@@

MRP allows users to specify priors that reflect knowledge of the variants and phenotypes under study. The prior on the scale of effect sizes can be used to denote which variants may have larger effect sizes. For instance, emerging empirical genetic studies have shown that within a gene, PTVs may have stronger effects than missense variants\cite{do2015exome}. This can be reflected by providing larger priors for PTVs.

@@

structure for protein XXX which was homologous to the \textit{ANGPTL7} protein

>>>>

\subsection*{Description of MRP}
In this section, we provide an overview of the MRP model comparison approach. Refer to \nameref{S1_Appendix} for a detailed description. MRP models GWAS summary statistics as being distributed according to one of two models. The null model is that the regression effect sizes obtained across all studies for a group of variants and a group of phenotypes is zero. The alternative model is that summary statistics are distributed according to a multivariate normal distribution with mean zero and covariance matrix described below. MRP compares the evidence for the null and alternative model using a Bayes Factor (BF) that quantifies the amount of evidence for each model as the ratio of the marginal likelihoods of the observed data under two models.

@@

MRP allows users to specify priors that reflect knowledge of the variants and phenotypes under study. For instance, we can define an independent effects model where each variant in the model may have different effect sizes. In this case, $\mathbf{R_{\textrm{var}}}$ is the identity matrix which reflects the assumption that the effect sizes of the variants are not correlated. This could be true for missense variants where some variants may impact important functional domains but others do not. We can also define a similar effects model by setting every value of $\mathbf{R_{\textrm{var}}}$ to $\sim 1$. This model assumes that all variants under consideration have the same effect sizes. This model may be appropriate for PTVs where each variant completely disrupts the function of the gene, leading to a gene knockout. The prior on the scale of effect sizes can also be used to denote which variants may have larger effect sizes. For instance, emerging empirical genetic studies have shown that within a gene, PTVs may have stronger effects than missense variants\cite{do2015exome}. This can be reflected by providing larger priors for PTVs.

@@
